\mutagen: Coverage-Guided, Property-Based Testing using Exhaustive Structure-Preserving Mutations

### 2022-03-09 papers/fse22/main.tex
General pass over the text

\begin{abstract}
Automatically synthesized random data generators are an appealing option when
using property-based testing.
%
There exists a variety of techniques that extract static information from the
codebase to produce random test cases.
%
Sadly, such techniques cannot derive generators capable of producing random
values satisfying complex invariants, often needed to test properties with
sparse preconditions.

@@

This work presents \mutagen, a CGPT framework that overcomes such limitations by
generating mutants \emph{exhaustively}.
%
Our tool incorporates heuristics to help minimizing scalability issues as well
as covering the search space in a principled manner.   
%
% schedule mutants based on their
% novelty and to minimize mutating certain subexpressions leading to scalability
% issues of exhaustiveness.
%of large enumeration types more
%than necessary.
%
Our evaluation show that \mutagen not only outperforms existing CGPT tools, but
also finds previously unknwown bugs in real-world software.
\end{abstract}

@@

%% PBT and generation of interesting data for free
%
Popularized by \quickcheck \cite{ClaessenH00}, Random Property-Based Testing
(RPBT) is a popular technique for finding bugs
\cite{ClaessenH00,hughes2003erlang, papadakis2011proper, bulwahn2012new,
  denes2014quickchick}.
%
% This is achieved by using randomly generated inputs to instantiate executable
It uses randomly generated inputs to instantiate executable properties encoding
the expected behavior of the system under test.
%
%% Exemplified by frameworks like Haskell's \quickcheck \cite{ClaessenH00} and
%the % Abundant ports of this tool in other programming languages %
%\cite{hughes2003erlang, papadakis2011proper, bulwahn2012new, %
%denes2014quickchick}
%
%% The most common PBT approach is to instantiate the testing properties using %
%randomly generated inputs.
%
% However, users of RPBT in general) are well aware that one of the biggest
% challenges while using RPBT is to provide suitable random data generators
% needed to instantiate the testing properties.
However, a limiting factor of RPBT is having to provide the random data
generators used to instantiate the testing properties. 
%
Writing highly-tuned generators capable of exercising every part of a complex
system on a reasonable basis can take several thousand person-hours of trial and
error \cite{lampropoulos2019coverage}.
%
Luckily, there exist several approaches to automatically synthesize random data
generators by reifying the static information present in the codebase, e.g.,
data type definitions, application public interfaces (APIs), etc.
\cite{GriecoCB16, DBLP:conf/haskell/MistaRH18, Mista2019GeneratingRS,
DuregardJW12, Lampropoulos2017, Bendkowski2017}.
%
These approaches, however, are unable to synthesize generators capable of
producing data satisfying complex invariants not easily derivable from the
codebase.
%
Randomly generating valid programs to test compilers is an example of this
limitation \cite{chen2020asurvey}, where developers are forced to put
substantial efforts in writing specialized generators by hand \cite{Palka11,
perenyi2020stack, yang2011finding}.
%
Moreover, automatically synthesized generators behave poorly when used to test
properties with sparse preconditions.
%
In this scenario, most tests are simply thrown away before reaching the main
components of the system under test.

@@

Unlike the generic bit-level mutators often used in the fuzzing community
\cite{zzuf,radamsa}, structure-preserving mutations specified at the data type
level can effectively produce syntactically valid mutants. 
%
% %
% High-level mutators avoid wasting time with generic low-level mutators that work
% at the bit level and know very little about data structures, thus producing
% syntactically broken mutants most of the time.
%
This grammar-aware technique has shown to be effective when fuzzing systems
accepting structurally complex inputs \cite{holler2012fuzzing, wang2019superion,
xsmith}.
%
In practice, most tools use existing grammars to tailor mutations to the input
structure used by the system under test. 
%
% Notably external grammars are not required in CGPT.
%
%% Tools like LangFuzz \cite{holler2012fuzzing}, Superion \cite{wang2019superion},
%% XSmith \cite{xsmith} use existing grammars to tailor the mutators to the
%% specific input structure used by the system under test.
%
% In CGPT, external grammars are not required.
%
In contast, CGPT relies on the data type information of the inputs to the
testing properties to derive specialized structure-preserving mutators directly
--- making strongly-typed programming languages an ideal match for this
technique.
%
% already provide a good source of information about how data can be mutated,
% e.g., by changing one data constructor by another.
%
%
% The data types used by the inputs of the testing properties already describe the
% structure of the random data one needs to mutate in a concrete manner, and
% Specialized mutators can be automatically derived directly from data types
% definitions ---
%

@@

%% Our solution
To tackle these limitations, we introduce \mutagen, a CGPT framework that
applies mutations \emph{exhaustively} (see \S \ref{sec:mutagen}).
%
%% implemented in Haskell
%% that also builds upon code instrumentation and type-preserving mutations
%
%
% the basic ideas behind \fuzzchick, i.e.,
%
%% \mutagen is introduced in detail in \S \ref{sec:mutagen}.
%
%% Exhaustive mutations
%% Unlike \fuzzchick, mutations in \mutagen are applied \emph{exhaustively}.
%
That is, given an interesting test case, our tool precomputes and schedules
every structure-preserving mutation that can be applied to it exactly once.
%
This approach has two inherent advantages. %when compared to \fuzzchick.
%
On one hand,
%  mutations do not rely on randomness to be generated.
%
every subexpression of the input test case is mutated on the same basis,
guaranteeing that no interesting transformation is omitted due to randomness.
%
%% On the other hand, scheduling mutations exhaustively eliminates the need for a
%% power schedule to assign a given mutation energy to each input test case.
%
On the other hand, scheduling mutations exhaustively eliminates the need for a
power schedule to assign a mutation budget to every interesting test case.

@@

\item \emph{Power schedule:}
%
it is unclear how well the heuristic that assigns a budget to each interesting
test case in CGPT performs in tandem with structure-preserving mutations derived
for an arbitrary data type.
%
On one hand, spending too much budget on not-so-interesting test cases might
prevent bugs from being discovered on a timely basis.
%
In turn, spending too little budget on interesting test cases might prevent bugs
from being discovered at all --- randomly generating the same initial test case
again later does not help, as it will no longer be considered interesting
according to historic trace information.
%
Finding a well behaved heuristic between these two scenario can be quite
challenging in the general case. 
%
\emph{Ideally, the scheduling mechanism should be as unbiased as possible.}

@@

In this section we describe the main ideas behind \mutagen, our CGPT tool
written in Haskell to tackle the limitations of CGPT described so far.
%
Notably, these ideas can be applied beyond Haskell and functional programming
languages in general.
%
%% Later, \S \ref{sec:heuristics} introduces heuristics that help finding bugs
%% faster in certain testing scenarios.

@@

\setlength{\intextsep}{10pt}%
\setlength{\columnsep}{10pt}%
\begin{wrapfigure}{r}{0.2\textwidth}
\vspace{-20pt}
\begin{verbatim}
mutate t = case t of 
  Leaf x -> [ 
    PURE (Branch def x def)
  ]
  Branch l x r -> [ 
    PURE l, PURE r, 
    PURE (Leaf x), 
    PURE (Branch l x l),
    PURE (Branch r x r),
    PURE (Branch r x l)
  ]
\end{verbatim}
\vspace{-10pt}
\caption{\label{fig:mutagen:mutator}\mutagen mutator for the \texttt{Tree} data
  type.}
\vspace{-10pt}
\end{wrapfigure}

To illustrate this, Fig. \ref{fig:mutagen:mutator} outlines a mutator for the
\texttt{Tree} data type.
%
This definition simply needs to enumerate mutants that transform the outermost
data constructor, hence no recursion is needed here.
%
Moreover, notice how a default value \texttt{def} used to fill the subtrees when
transforming a leaf into a branch.
%
In practice, \texttt{def} corresponds to the simplest expression we can
construct for the mutant to be type-correct.
%
In our example, the default value of a \texttt{Tree} is a leaf containing the
smallest value of the tree payload type, e.g., in the case of trees with
integers \texttt{def} is defined as \texttt{Leaf 0}.
%
%% \texttt{Leaf def}, where
%% the inner \texttt{def} is the default value of the payload.%
(Our implementation uses the type class system to abstract this complexity
away.)
%
Using a small default value, as oppossed to a randomly generated one (as done by
the original CGPT), is also inspired by exhaustive bounded testing tools, and
avoids introducing unnecessary randomness when ``growing'' a data constructor
into another.

@@

\noindent Our tool automatically synthesizes \texttt{mutate} to follow the
equation:

@@

\noindent And $\mathtt{mut_s}$ builds all the possible mutations that swap the
root data constructor with a (possibly) different one, reusing (or defaulting)
compatible subexpresions to instantiate the new data constructor:

\vspace{-10pt}
\begin{flalign*}
\mathtt{mut_{s}}(C_i \, x_1 \, x_2 \, \mydots) \!=\! 
  \Vast\{ 
    \mathtt{PURE}(C_j \, x_1^\prime \, x_2^\prime \, \mydots)
    \,\Vast|\,
    \begin{aligned}
      & C_j^T \!\in\! \{ C_1, C_2, \mydots \} \\
      & x_k^\prime \!\in\! \overline{\mathtt{filter}}(t_k^j, \{x_1, \, x_2, \, \mydots \}) \\  
      & C_j \, x_1^\prime \, x_2^\prime \, \mydots \;\neq\; C_i \, x_1 \, x_2 \, \mydots
    \end{aligned}
  \Vast\} &&
\end{flalign*}
\vspace{-7pt}

\noindent Where the helper function $\mathtt{filter}$ simply returns the subset
of the input values $X$ that match a target type $t$:

@@

\noindent And the function $\overline{\mathtt{filter}}$ is a wrapper over
$\mathtt{filter}$ that returns the default value of the target type $t$ if there
are no compatible subexpressions available to reuse:

@@

With the exhaustive mutation mechanism in place, we can now introduce the base
testing loop used by \mutagen --- outlined in Algorithm \ref{algo:mutagen:loop}.
%
Like in CGPT, we use two queues, QValid and QDiscarded to keep valid and
discarded mutable test cases, respectively.
%
%% The testing loop enqueues and retrieves interesting test cases from them before
%% falling back to random generation.

@@

Finally, the seed selection algorithm (Algorithm \ref{algo:mutagen:pick}) picks
the next test case using the same criteria as CGPT, prioritizing valid test
cases over discarded ones, falling back to random generation when both queues
are empty.
%
This function simply picks the next mutated test case from the current
precomputed batch, jumping to the next batch in line when the current one it
becomes empty.

@@

@@

Every time we insert a new execution path into the internal trace log (TLog), we
calculate the number of \emph{new nodes} that were executed, as well as the
\emph{branching depth} where they got inserted.
%
The former is used to distinguish interesting test cases (whether or not new
nodes were inserted), whereas the latter is used by this heuristic to prioritize
some test cases over others.
%
Fig. \ref{fig:mutagen:tracelog} illustrates this idea, inserting two execution
traces (one after another) into a trace log that initially contains a single
execution path.
%
The second insertion (with trace $1 \rightarrow 2 \rightarrow 6 \rightarrow 7$)
branches at a shallower depth than the first one (2 vs. 3), and so its
corresponding test case is given a higher priority.

Using this mechanism, we can modify \mutagen's base testing loop by replacing
each mutation queue with a priority queue indexed by the branching depth of each
new execution.
%
These changes are illustrated in Algorithm \ref{algo:mutagen:lifo}.
%
Statements in {\color{red} red} indicate important changes to the base
algorithm, whereas ellipses denote parts of the code that are not relevant for
the point being made.

@@

As introduced in \S \ref{sec:mutagen}, our tool is parameterized by the number
of mutated test cases to be generated for each subexpression of the input test
case associated to random mutant.
%
But, how many random test cases should we sample? A single one? A few tenths? A
few hundred?
%% %
%% Using too little can put finding bugs at risk.
%% %
%% For instance, when the system under test branches based on numeric values, we
%% should make sure that we set enough random mutations to test each branch on a
%% reasonable basis.
%% %
%% Using too many, on the other hand, can degrade the performance of the testing
%% loop, as it will spend too much time producing uninteresting mutations.
%% %
%% This can happen for instance if the subexpressions defined using random mutants
%% are only used as payloads, and their value does not affect the execution in any
%% way.
%
Answering this question precisely can be challenging, so this second heuristic
aims to alleviate the problem.

We found that, the smaller the number of random mutations we initially set, the
easier it is for the trace log that records executions to get saturated, i.e.,
when interesting test cases stop getting discovered or are discovered very
sporadically.
%
We realized that we can use this information to automatically optimize the
number of random mutations used by our tool.
%
This idea is described in Algorithm \ref{algo:mutagen:reset}.
%
The process is as follows:
%
\begin{inparaenum}
  \item we start the testing loop with the number of random mutations set to
    one, and
  %
  \item each time we find that a test is not interesting (i.e. boring), we
    increment a counter.
  %
  Then,
  \item if we have not produced any interesting test case after a certain number
    of tests (1000 tests seems to be a reasonable threshold in practice), we
    increment the number of random mutations and the threshold by twice the
    current amount.
    %
    Additionally, each time this happens we reset the trace log, so interesting
    test cases found on a previous iteration can be found and enqueued for
    mutation again with a higher effort dedicated to sampling mutants.
\end{inparaenum}

Notably, this heuristic can be useful when the execution of the system under
test depends heavily on some invariant encoded by a subexpression of its input
that get transformed using a random mutant (e.g, the number of pixels declared
by the header of an image matching the size of its actual payload).
%
There, starting with a single random mutation will quickly saturate the trace
space with discarded (invalid) tests, and this heuristic will continuously
increase the random mutations parameter until the random mutant is given enough
chances to randomly generate a value that satisfies the required invariant and
makes the overall test case valid.

@@

\noindent Control flow is achieved using the \texttt{Call} and \texttt{Ret}
instructions, which let the program jump back and forth within the instruction
memory using specially labeled values in the stack representing memory
addresses.
%
The argument of the \texttt{Push} instruction represents a value to be inserted
in the stack, whereas the argument of the \texttt{Call} instruction encodes the
number of elements of the stack to be treated as arguments.
%
Then, programs are simply modeled as sequences of instructions.
%
Finally, machine states are modeled using a 4-tuple \textit{(pc, stk, m, i)}
representing a particular configuration of the program counter, the stack, and
the data- and instruction memories, respectively.
%
To preserve space, we encourage the reader to refer to the work of
\citeauthor{hritcu2013testing} \citeyear{hritcu2013testing,
  hrictcu2016testing} and \citeauthor{lampropoulos2019coverage} for more details
about the implementation and semantics of this case study.

@@

The IFC rules enforced by this machine are encoded in a table indexed by its
different instructions.
%
This table stores the dynamic check that the machine performs to enforce the IFC
policy before running each instruction, along with the corresponding resulting
security labels of the program counter and the instruction result.
%
For instance, to execute the \texttt{Store} instruction (which stores a value in
a memory pointer) the machine checks that both the labels of the program counter
and the pointer together can flow to the label of the destination memory cell.
%
If this condition is not met, the machine halts to indicate a violation of the
IFC policy.
%
After this check, the machine overwrites the value at the destination cell and
updates its label with the maximum sensibility of the involved labels.
%
In the rule table, this looks as follows:

@@

Having sorted these issues out, we mechanically injected 10 new bugs in the
validator as well as 5 new bugs in the interpreter of this engine (see Table
\ref{table:wasm:injectedbugs}).
%
These bugs either
%
\begin{inparaenum}
\item remove an existing integrity check (to weaken the WebAssembly
  type-system/validator); or
\item simulate a copy-paste bug \cite{CPMiner},%
  \footnote{This kind of bugs was inspired by the real bug \#1 we found prior to
    this step.}
  %
  replacing the implementation of an instruction with a compatible one
  (e.g., \texttt{i32.add} by \texttt{i32.sub}).
\end{inparaenum}

@@

This testing property instantiates a module stub \texttt{m} with the input
WebAssembly function (\texttt{fun}) and its type signature (\texttt{ty}).
%
Then, it invokes the function \texttt{f} of the module \texttt{m} both on
\textit{haskell-wasm} and the reference interpreter with the provided arguments
(\texttt{args}).
%
Finally, the property asserts whether their results are equivalent.%
%
\footnote{We additionally set a short timeout to discard potentially diverging
  programs.}
%
Interestingly, equivalence in does not imply equality.
%
Non-deterministic operations in WebAssembly like NaN reinterpretations can
produce different equivalent results (as exposed by the discrepancy \#5 in Table
\ref{table:wasm:bugs}), and our equivalence relation needs to take that into
account.
%

Using this testing property directly might not sound wise, as randomly generated
lists of input arguments will very unlikely match the type signature of randomly
generated functions.
%
However, it lets us test what happens when programs are not properly invoked,
and it quickly discovered the previously unknown bug \#3 in
\textit{haskell-wasm} mentioned above.
%
Having fixed this issue, we define a specialized version of
\texttt{prop\_interpreter} that fixes the type of the generated function to take
two arguments (of type \texttt{I32} and \texttt{F32}) and return an \texttt{I32}
as a result:

@@

\noindent This property lets us generate functions of a fixed type and invoke
them with randomly generated input of the expected types.
%
%% In our experiments (presented in \S \ref{sec:evaluation}),
%
We use this property to find all the bugs injected into \textit{haskell-wasm}'s
interpreter in the next section.
%
%% As before, we remark that a complete test suite should at least include
%% different variants of this property testing functions of several different
%% types, as well as properties testing multiple functions simultaneously.

@@

We performed all experiments in a workstation with 32GB of RAM and an Intel Core
i7-8700 CPU running at 3.20GHz.
%
Each experiment was repeated 30 times except for the ones testing the
WebAssembly interpreter, which were run 10 times due to time constraints.
%
Moreover, we followed the approach taken by
\citeauthor{lampropoulos2019coverage} and collected the mean-time-to-failure
(MTTF) of each bug, i.e., how quickly a bug can be found in wall clock time.
%
We used a one-hour timeout to stop the execution of both tools if they have not
yet found a counterexample.
%
In addition, we collected the failure rate (FR) observed for each bug, i.e. the
proportion of times each tool finds each bug within the one-hour testing budget.
%
%% We found this metric crucial to be analyzed when replicating \fuzzchick's
%% results, as opposed to just paying attention at the MTTF.

>>>>

\begin{abstract}
Automatically synthesized random data generators are an appealing option when
using property-based testing.
%
There exists a variety of techniques that extract static information from the
codebase to produce random test cases.
%
Sadly, such techniques cannot derive generators capable of producing random
values satisfying complex invariants often needed to test properties with sparse
preconditions.

@@

This work presents \mutagen, a CGPT framework that overcomes such limitations by
generating mutants \emph{exhaustively}.
%
Our tool incorporates heuristics that help to minimize scalability issues as well
as covering the search space in a principled manner.   
%
% schedule mutants based on their
% novelty and to minimize mutating certain subexpressions leading to scalability
% issues of exhaustiveness.
%of large enumeration types more
%than necessary.
%
Our evaluation shows that \mutagen not only outperforms existing CGPT tools but
also finds previously unknown bugs in real-world software.
\end{abstract}

@@

%% PBT and generation of interesting data for free
%
Popularized by \quickcheck \cite{ClaessenH00}, Random Property-Based Testing
(RPBT) is a popular technique for finding bugs
\cite{ClaessenH00,hughes2003erlang, papadakis2011proper, bulwahn2012new,
  denes2014quickchick}.
%
% This is achieved by using randomly generated inputs to instantiate executable
It uses randomly generated inputs to instantiate executable properties encoding
the expected behavior of the system under test.
%
%% Exemplified by frameworks like Haskell's \quickcheck \cite{ClaessenH00} and
%the % Abundant ports of this tool in other programming languages %
%\cite{hughes2003erlang, papadakis2011proper, bulwahn2012new, %
%denes2014quickchick}
%
%% The most common PBT approach is to instantiate the testing properties using %
%randomly generated inputs.
%
% However, users of RPBT in general) are well aware that one of the biggest
% challenges while using RPBT is to provide suitable random data generators
% needed to instantiate the testing properties.
However, a limiting factor of RPBT is the need for the random data generators
used to instantiate the testing properties. 
%
Writing highly-tuned generators capable of exercising every part of a complex
system on a reasonable basis can take several thousand person-hours of trial and
error \cite{lampropoulos2019coverage}.
%
Luckily, there exist several approaches that automatically synthesize random
data generators by reifying the static information present in the codebase,
e.g., data type definitions, application public interfaces (APIs), etc.
\cite{GriecoCB16, DBLP:conf/haskell/MistaRH18, Mista2019GeneratingRS,
DuregardJW12, Lampropoulos2017, Bendkowski2017}.
%
These approaches, however, are unable to synthesize generators capable of
producing data satisfying complex invariants not easily derivable from the
codebase.
%
Randomly generating valid programs to test compilers is an example of this
limitation \cite{chen2020asurvey}, where developers are forced to put
substantial efforts into writing specialized generators by hand \cite{Palka11,
perenyi2020stack, yang2011finding}.
%
Moreover, automatically synthesized generators behave poorly when used to test
properties with sparse preconditions.
%
In this scenario, most tests are simply thrown away before reaching the main
components of the system under test.

@@

Unlike the generic bit-level mutators often used in the fuzzing community
\cite{zzuf,radamsa}, structure-preserving mutations specified at the data type
level can effectively produce syntactically valid mutants. 
%
% %
% High-level mutators avoid wasting time with generic low-level mutators that work
% at the bit level and know very little about data structures, thus producing
% syntactically broken mutants most of the time.
%
This grammar-aware technique has shown to be effective when fuzzing systems
accepting structurally complex inputs \cite{holler2012fuzzing, wang2019superion,
xsmith}.
%
In practice, most tools use existing grammars to tailor mutations to the input
structure used by the system under test. 
%
% Notably external grammars are not required in CGPT.
%
%% Tools like LangFuzz \cite{holler2012fuzzing}, Superion \cite{wang2019superion},
%% XSmith \cite{xsmith} use existing grammars to tailor the mutators to the
%% specific input structure used by the system under test.
%
% In CGPT, external grammars are not required.
%
In contrast, CGPT relies on the data type information of the inputs to the
testing properties to derive specialized structure-preserving mutators directly
--- making strongly-typed programming languages an ideal match for this
technique.
%
% already provide a good source of information about how data can be mutated,
% e.g., by changing one data constructor by another.
%
%
% The data types used by the inputs of the testing properties already describe the
% structure of the random data one needs to mutate in a concrete manner, and
% Specialized mutators can be automatically derived directly from data types
% definitions ---
%

@@

%% Our solution
To tackle these limitations, we introduce \mutagen, a CGPT framework that
applies mutations \emph{exhaustively} (see \S \ref{sec:mutagen}).
%
%% implemented in Haskell
%% that also builds upon code instrumentation and type-preserving mutations
%
%
% the basic ideas behind \fuzzchick, i.e.,
%
%% \mutagen is introduced in detail in \S \ref{sec:mutagen}.
%
%% Exhaustive mutations
%% Unlike \fuzzchick, mutations in \mutagen are applied \emph{exhaustively}.
%
That is, given an interesting test case, our tool precomputes and schedules
every structure-preserving mutation that can be applied to it to be evaluated
exactly once.
%
This approach has two inherent advantages. %when compared to \fuzzchick.
%
Firstly,
%  mutations do not rely on randomness to be generated.
%
every subexpression of the input test case is mutated on the same basis,
guaranteeing that no interesting transformation is omitted due to randomness.
%
%% On the other hand, scheduling mutations exhaustively eliminates the need for a
%% power schedule to assign a given mutation energy to each input test case.
%
Moreover, scheduling mutations exhaustively eliminates the need for a
power schedule to assign a mutation budget to every interesting test case.

@@

\item \emph{Power schedule:}
%
it is unclear how well the heuristic that assigns a budget to each interesting
test case in CGPT performs in tandem with structure-preserving mutations derived
for an arbitrary data type.
%
On one hand, spending too much budget on not-so-interesting test cases might
prevent bugs from being discovered on a timely basis.
%
In turn, spending too little budget on interesting test cases might prevent bugs
from being discovered at all --- randomly generating the same initial test case
again later does not help, as it will no longer be considered interesting
according to historic trace information.
%
Finding a well-behaved heuristic that lies between these two scenarios can be
quite challenging in the general case. 
%
\emph{Ideally, the scheduling mechanism should be as unbiased as possible.}

@@

In this section we describe the main ideas behind \mutagen, our revised CGPT
tool written in Haskell that tackles the limitations of CGPT described so far.
%
Notably, these ideas can be applied beyond Haskell and functional programming
languages in general.
%
%% Later, \S \ref{sec:heuristics} introduces heuristics that help finding bugs
%% faster in certain testing scenarios.

@@

\setlength{\intextsep}{10pt}%
\setlength{\columnsep}{10pt}%
\begin{wrapfigure}{r}{0.2\textwidth}
\vspace{-20pt}
\begin{verbatim}
mutate t = case t of 
  Leaf x -> [ 
    PURE (Branch def x def)
  ]
  Branch l x r -> [ 
    PURE l, PURE r, 
    PURE (Leaf x), 
    PURE (Branch l x l),
    PURE (Branch r x r),
    PURE (Branch r x l)
  ]
\end{verbatim}
\vspace{-10pt}
\caption{\label{fig:mutagen:mutator}\mutagen mutator for the \texttt{Tree} data
  type.}
\vspace{-20pt}
\end{wrapfigure}

To illustrate this, Fig. \ref{fig:mutagen:mutator} outlines a mutator for the
\texttt{Tree} data type.
%
This definition simply needs to enumerate mutants that transform the outermost
data constructor, hence no recursion is needed here.
%
Moreover, notice how a default value \texttt{def} used to fill the subtrees when
transforming a leaf into a branch.
%
In practice, \texttt{def} corresponds to the simplest expression we can
construct for the mutant to be type-correct.
%
In our example, the default value of a \texttt{Tree} is a leaf containing the
smallest value of the tree payload type, e.g., in the case of trees with
integers \texttt{def} is defined as \texttt{Leaf 0}.
%
%% \texttt{Leaf def}, where
%% the inner \texttt{def} is the default value of the payload.%
(Our implementation uses the type class system to abstract this complexity
away.)
%
Using a small default value, as opposed to a randomly generated one (as done by
the original CGPT), is also inspired by exhaustive bounded testing tools, and
avoids introducing unnecessary randomness when ``growing'' a data constructor
into another.

@@

\noindent Our tool automatically synthesizes \texttt{mutate} by pattern matching
on the root data constructor of the input. 
%
On each case, \texttt{mutate} follows:

@@

\noindent Then, $\mathtt{mut_s}$ builds all the possible mutations that swap the
root data constructor with a (possibly) different one, reusing (or defaulting)
compatible subexpresions to instantiate the new data constructor:

\vspace{-10pt}
\begin{flalign*}
\mathtt{mut_{s}}(C_i \, x_1 \, x_2 \, \mydots) \!=\! 
  \Vast\{ 
    \mathtt{PURE}(C_j \, x_1^\prime \, x_2^\prime \, \mydots)
    \,\Vast|\,
    \begin{aligned}
      & C_j \!\in\! \{ C_1, C_2, \mydots \} \\
      & x_k^\prime \!\in\! \overline{\mathtt{filter}}(t_k^j, \{x_1, \, x_2, \, \mydots \}) \\  
      & C_j \, x_1^\prime \, x_2^\prime \, \mydots \;\neq\; C_i \, x_1 \, x_2 \, \mydots
    \end{aligned}
  \Vast\} &&
\end{flalign*}
\vspace{-7pt}

\noindent In the previous definitions, the helper function $\mathtt{filter}$
simply returns the subset of the input values $X$ that match a target type $t$:

@@

\noindent Whereas the function $\overline{\mathtt{filter}}$ is a wrapper over
$\mathtt{filter}$ that returns the default value of the target type $t$ if there
are no compatible subexpressions available to reuse:

@@

With the exhaustive mutation mechanism in place, we can now introduce the base
testing loop used by \mutagen --- outlined in Algorithm \ref{algo:mutagen:loop}.
%
Like in CGPT, we use two queues, QValid and QDiscarded to keep valid and
discarded interesting test cases, respectively.
%
%% The testing loop enqueues and retrieves interesting test cases from them before
%% falling back to random generation.

@@

Then, the seed selection algorithm (Algorithm \ref{algo:mutagen:pick}) picks
the next test case using the same criteria as CGPT, prioritizing valid test
cases over discarded ones, falling back to random generation when both queues
are empty.
%
This function simply picks the next mutated test case from the current
precomputed batch, jumping to the next batch in line when the current one it
becomes empty.

@@

Every time we insert a new execution path into the internal trace log (TLog), we
calculate the number of \emph{new nodes} that were executed, as well as the
\emph{branching depth} where they got inserted.
%
The former is used to distinguish interesting test cases (whether or not new
nodes were inserted), whereas the latter is used by this heuristic to prioritize
some test cases over others.
%
Fig. \ref{fig:mutagen:tracelog} illustrates this idea, inserting two execution
traces (one after another) into a trace log that initially contains a single
execution path.
%
Both executions are considered interesting, but the second one (with trace $1
\rightarrow 2 \rightarrow 6 \rightarrow 7$) branches at a shallower depth than
the first one (2 vs. 3), and so its corresponding test case is given a higher
priority.

Using this mechanism, we can modify \mutagen's base testing loop by replacing
each mutation queue with a priority queue indexed by the branching depth of each
new execution.
%
These changes are illustrated in Algorithm \ref{algo:mutagen:lifo}.
%
Statements in {\color{red} red} indicate important changes to the base
algorithm, whereas ellipses denote unchaged parts of the code that are not
relevant for the point being made.

@@

@@

As introduced in \S \ref{sec:mutagen}, our tool is parameterized by the number
of mutated test cases to be generated for each subexpression of the input test
case associated to a random mutation.
%
But, how many random test cases should we sample? A single one? A few tenths? A
few hundred?
%% %
%% Using too little can put finding bugs at risk.
%% %
%% For instance, when the system under test branches based on numeric values, we
%% should make sure that we set enough random mutations to test each branch on a
%% reasonable basis.
%% %
%% Using too many, on the other hand, can degrade the performance of the testing
%% loop, as it will spend too much time producing uninteresting mutations.
%% %
%% This can happen for instance if the subexpressions defined using random mutants
%% are only used as payloads, and their value does not affect the execution in any
%% way.
%
Answering this question precisely can be challenging, so this second heuristic
aims to alleviate the problem.

We found that, the smaller the number of random mutations we initially set, the
easier it is for the trace log that records executions to get saturated, i.e.,
when interesting test cases stop getting discovered or are discovered very
sporadically.
%
We realized that we can use this information to automatically optimize the
number of random mutations used by our tool.
%
This idea is described in Algorithm \ref{algo:mutagen:reset}.
%
The process is as follows:
%
\begin{inparaenum}
  \item we start the testing loop with the number of random mutations set to
    one, and
  %
  \item each time we find that a test is not interesting (i.e. boring), we
    increment a counter.
  %
  Then,
  \item if we have not produced any interesting test case after a certain number
    of tests (1000 tests seems to be a reasonable threshold in practice), we
    increment the number of random mutations and the threshold by twice the
    current amount.
    %
    Additionally, each time this happens we reset the trace log, so interesting
    test cases found on a previous iteration can be found and enqueued for
    mutation again with a higher effort dedicated to random mutations.
\end{inparaenum}

Notably, this heuristic can be useful when the execution of the system under
test depends heavily on some invariant encoded by a subexpression of its input
that get transformed using a random mutation (e.g, the number of pixels declared
by the header of an image matching the size of its actual payload).
%
There, starting with a single random mutation will quickly saturate the trace
space with discarded (invalid) tests, and this heuristic will continuously
increase the random mutations parameter until the random mutant is given enough
chances to randomly generate a value that satisfies the required invariant and
makes the overall test case valid.

@@

\noindent Control flow is achieved using the \texttt{Call} and \texttt{Ret}
instructions, which let the program jump back and forth within the instruction
memory using specially labeled values in the stack representing memory
addresses.
%
The argument of the \texttt{Push} instruction represents a value to be inserted
in the stack, whereas the argument of the \texttt{Call} instruction encodes the
number of elements of the stack to be treated as arguments.
%
Then, programs are simply modeled as sequences of instructions.
%
Finally, machine states are modeled using a 4-tuple \textit{(pc, stk, m, i)}
representing a particular configuration of the program counter, the stack, and
the data- and instruction memories, respectively.
%
To preserve space, we encourage the reader to refer to the work of
\citeauthor{hritcu2013testing} \cite{hritcu2013testing, hrictcu2016testing} and
\citeauthor{lampropoulos2019coverage} \cite{lampropoulos2019coverage} for more
details about the implementation and semantics of this case study.

@@

The IFC rules enforced by this machine are encoded in a table indexed by its
different instructions.
%
This table stores the dynamic check that the machine performs to enforce the IFC
policy before running each instruction, along with the resulting security labels
corresponding to the program counter and the instruction result.
%
For instance, to execute the \texttt{Store} instruction (which stores a value in
a memory pointer) the machine checks that both the labels of the program counter
and the pointer together can flow to the label of the destination memory cell.
%
If this condition is not met, the machine halts to indicate a violation of the
IFC policy.
%
After this check, the machine overwrites the value at the destination cell and
updates its label with the maximum sensibility of the involved labels.
%
In the rule table, this looks as follows:

@@

Having sorted these issues out, we mechanically injected 10 new bugs in the
validator as well as 5 new bugs in the interpreter of this engine (see Table
\ref{table:wasm:injectedbugs}).
%
These bugs either
%
\begin{inparaenum}
\item remove an existing integrity check (to weaken the WebAssembly
  type-system/validator); or
\item simulate a copy-paste bug \cite{chou2001empirical},
  %
  replacing the implementation of an instruction with a compatible one
  (e.g., \texttt{i32.add} by \texttt{i32.sub}).%
  \footnote{This kind of bugs was inspired by the real bug \#1 we found prior to
    this step.}
\end{inparaenum}

@@

This testing property instantiates a module stub \texttt{m} with the input
WebAssembly function (\texttt{fun}) and its type signature (\texttt{ty}).
%
Then, it invokes the function \texttt{f} of the module \texttt{m} both on
\textit{haskell-wasm} and the reference interpreter with the provided arguments
(\texttt{args}).
%
Finally, the property asserts whether their results are equivalent.%
%
\footnote{We additionally set a short timeout to discard potentially diverging
  programs.}
%
Interestingly, equivalence does not imply equality.
%
Non-deterministic operations in WebAssembly like NaN reinterpretations can
produce different equivalent results (as exposed by the discrepancy \#5 in Table
\ref{table:wasm:bugs}), and our equivalence relation needs to take that into
account.
%

Using this testing property directly might not sound wise, as randomly generated
lists of input arguments will very unlikely match the type signature of randomly
generated functions.
%
However, it lets us test what happens when programs are not properly invoked,
and it quickly discovered the previously unknown bug \#3 in
\textit{haskell-wasm} mentioned above.
%
Having fixed this issue, we define a specialized version of
\texttt{prop\_interpreter} that fixes the type of the generated function to take
two arguments \texttt{x} and \texttt{y} (of type \texttt{I32} and \texttt{F32},
respectively) and return an \texttt{I32} as a result:

@@

\noindent This property lets us generate functions of a fixed type and invoke
them with randomly generated inputs of the expected types.
%
%% In our experiments (presented in \S \ref{sec:evaluation}),
%
We use this property to find all the bugs injected into \textit{haskell-wasm}'s
interpreter in the next section.
%
%% As before, we remark that a complete test suite should at least include
%% different variants of this property testing functions of several different
%% types, as well as properties testing multiple functions simultaneously.

@@

We performed all experiments in a workstation with 32GB of RAM and an Intel Core
i7-8700 CPU running at 3.20GHz.
%
Each experiment was repeated 30 times except for the ones testing the
{Web-Assembly} interpreter, which were run 10 times due to time constraints.
%
Moreover, we followed the approach taken by
\citeauthor{lampropoulos2019coverage} and collected the mean-time-to-failure
(MTTF) of each bug, i.e., how quickly a bug can be found in wall clock time.
%
We used a one-hour timeout to stop the execution of both tools if they had not
yet found a counterexample.
%
In addition, we collected the failure rate (FR) observed for each bug, i.e. the
proportion of times each tool finds each bug within the one-hour testing budget.
%
%% We found this metric crucial to be analyzed when replicating \fuzzchick's
%% results, as opposed to just paying attention at the MTTF.
